Automated Residual Plot Assessment with the R Package autovi and the Shiny App autovi.web

### 2025-09-04 autovi_paper.tex
Fix the format for packages/software

\begin{abstract}
Visual assessment of residual plots is a common approach for diagnosing
linear models, but it relies on manual evaluation, which does not scale
well and can lead to inconsistent decisions across analysts. The lineup
protocol, which embeds the observed plot among null plots, can reduce
subjectivity but requires even more human effort. In today's data-driven
world, such tasks are well-suited for automation. We present a new R
package that uses a computer vision model to automate the evaluation of
residual plots. An accompanying Shiny app is provided for ease of use.
Given a sample of residuals, the model predicts a visual signal strength
(VSS) and offers supporting information to help analysts assess model
fit.
\end{abstract}

@@

Regression analysis is a widely used statistical modeling technique for
data in many fields. There is a vast array of software for conducting
regression modeling and generating diagnostics. The package
\texttt{lmtest} \citep{lmtest} provides a suite of conventional tests,
while the \texttt{stats} package \citep{stats} offers standard
diagnostic plots such as residuals vs.~fitted values, quantile-quantile
(Q-Q) plots, and residuals vs.~leverage plots. Additional packages like
\texttt{jtools} \citep{jtools}, \texttt{olsrr} \citep{olsrr},
\texttt{rockchalk} \citep{rockchalk}, and \texttt{ggResidpanel}
\citep{ggresidpanel} deliver similar graphical diagnostics, often with
enhanced aesthetics or interactive features. These tools collectively
produce the core diagnostic plots outlined in the classical text by
\citet{cook1982residuals}. The \texttt{ecostats} package
\citep{warton_global_2023} extends these diagnostics by incorporating
simulation envelopes into residual plots. Meanwhile, \texttt{DHARMa}
\citep{dharma} compares empirical quantiles (0.25, 0.5, and 0.75) of
scaled residuals to their theoretical counterparts, with a strong focus
on identifying model violations such as heteroscedasticity, misspecified
functional forms, and issues specific to generalized linear and
mixed-effect models, like over/under-dispersion. It also provides
conventional test annotations to reduce the risk of misinterpretation.

However, relying solely on subjective assessments of these plots can
lead to issues such as over-interpreting random patterns as model
violations. \citet{li2024plot} demonstrated that visual inference
methods, particularly those using the lineup protocol
\citep{buja2009statistical}, offer more practical and reliable
assessments of residual patterns than conventional tests, as they are
less sensitive to minor departures. Packages such as \texttt{nullabor}
\citep{nullabor}, \texttt{HLMdiag} \citep{loy2014hlmdiag}, and
\texttt{regressinator} \citep{regressinator} support this approach by
enabling users to compare observed residual plots with plots generated
under null hypothesis, thereby helping to quantify the significance of
any detected patterns.

@@

To make the statistical testing procedure and trained computer vision
model widely accessible, we developed the R package \texttt{autovi}
along with a companion web interface, \texttt{autovi.web}, which allows
users to automatically assess their residual plots using the trained
computer vision model.

The remainder of this paper is structured as follows:
Section~\ref{sec-vss-desc} introduces the definition and computation of
visual signal strength. Section~\ref{sec-autovi} provides a detailed
documentation of the \texttt{autovi} package, including its usage and
infrastructure. Section~\ref{sec-autovi-web} focuses on the
\texttt{autovi.web} interface, describing its design and usage, along
with illustrative examples. Finally, Section~\ref{sec-autovi-conclusion}
presents the main conclusions of this work.

>>>>

\begin{abstract}
Visual assessment of residual plots is a common approach for diagnosing
linear models, but it relies on manual evaluation, which does not scale
well and can lead to inconsistent decisions across analysts. The lineup
protocol, which embeds the observed plot among null plots, can reduce
subjectivity but requires even more human effort. In today's data-driven
world, such tasks are well-suited for automation. We present a new
\textsf{R} package that uses a computer vision model to automate the
evaluation of residual plots. An accompanying \textsf{Shiny} app is
provided for ease of use. Given a sample of residuals, the model
predicts a visual signal strength (VSS) and offers supporting
information to help analysts assess model fit.
\end{abstract}

@@

Regression analysis is a widely used statistical modeling technique for
data in many fields. There is a vast array of software for conducting
regression modeling and generating diagnostics. The package
\textsf{lmtest} \citep{lmtest} provides a suite of conventional tests,
while the \textsf{stats} package \citep{stats} offers standard
diagnostic plots such as residuals vs.~fitted values, quantile-quantile
(Q-Q) plots, and residuals vs.~leverage plots. Additional packages like
\textsf{jtools} \citep{jtools}, \textsf{olsrr} \citep{olsrr},
\textsf{rockchalk} \citep{rockchalk}, and \textsf{ggResidpanel}
\citep{ggresidpanel} deliver similar graphical diagnostics, often with
enhanced aesthetics or interactive features. These tools collectively
produce the core diagnostic plots outlined in the classical text by
\citet{cook1982residuals}. The \textsf{ecostats} package
\citep{warton_global_2023} extends these diagnostics by incorporating
simulation envelopes into residual plots. Meanwhile, \textsf{DHARMa}
\citep{dharma} compares empirical quantiles (0.25, 0.5, and 0.75) of
scaled residuals to their theoretical counterparts, with a strong focus
on identifying model violations such as heteroscedasticity, misspecified
functional forms, and issues specific to generalized linear and
mixed-effect models, like over/under-dispersion. It also provides
conventional test annotations to reduce the risk of misinterpretation.

However, relying solely on subjective assessments of these plots can
lead to issues such as over-interpreting random patterns as model
violations. \citet{li2024plot} demonstrated that visual inference
methods, particularly those using the lineup protocol
\citep{buja2009statistical}, offer more practical and reliable
assessments of residual patterns than conventional tests, as they are
less sensitive to minor departures. Packages such as \textsf{nullabor}
\citep{nullabor}, \textsf{HLMdiag} \citep{loy2014hlmdiag}, and
\textsf{regressinator} \citep{regressinator} support this approach by
enabling users to compare observed residual plots with plots generated
under null hypothesis, thereby helping to quantify the significance of
any detected patterns.

@@

To make the statistical testing procedure and trained computer vision
model widely accessible, we developed the R package \textsf{autovi}
along with a companion web interface, \textsf{autovi.web}, which allows
users to automatically assess their residual plots using the trained
computer vision model.

The remainder of this paper is structured as follows:
Section~\ref{sec-vss-desc} introduces the definition and computation of
visual signal strength. Section~\ref{sec-null-and-boot-desc} expands on
the computation of the null and bootstrapped residuals.
Section~\ref{sec-autovi} provides a detailed documentation of the
\textsf{autovi} package, including its usage and infrastructure.
Section~\ref{sec-autovi-web} focuses on the \textsf{autovi.web}
interface, describing its design and usage, along with illustrative
examples. Finally, Section~\ref{sec-autovi-conclusion} presents the main
conclusions of this work.
